---
title: "NFL 4th Down Win Probability Analysis"
author: "Neil Dorsey"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This analysis builds a win probability model to evaluate the decision-making of NFL coaches on 4th down plays using simulated data.

## Data Loading

```{r}
library(tidyverse)
data <- read.csv("nfl_4th_down_decision_data.csv")
head(data)
```

## Data Preparation

```{r}
data <- data %>%
  mutate(
    decision_factor = as.factor(decision),
    result_factor = as.factor(result),
    yard_position = ifelse(grepl("OWN", yard_line), "Own", "Opp"),
    yard_number = as.numeric(gsub("[^0-9]", "", yard_line))
  )
```

## Model Building

```{r}
# Predict optimal win probability based on situation
model <- lm(wp_optimal ~ to_go + yard_number + decision_factor, data = data)
summary(model)
```

## Evaluation of Coaching Decisions

```{r}
data <- data %>%
  mutate(
    wp_delta = round(wp_optimal - wp_actual, 3),
    missed_opportunity = ifelse(wp_delta > 0.05 & decision != "Go for it", TRUE, FALSE)
  )

# Summary
missed_summary <- data %>% filter(missed_opportunity == TRUE)
nrow(missed_summary)
head(missed_summary)
```

## Insights

- Coaches left potential win probability on the table in multiple scenarios.
- Model identified when “Go for it” had a higher expected value.
- Further tuning and validation could improve predictive accuracy.

## Conclusion

This model provides a starting point to better understand and optimize 4th down decisions using data.
